---
title: "part 4 Data Manipulation in R"
author: "Ilir_Sheraj"
date: "7/15/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Packages:

{tidyverse}

{magrittr}: Offers pipe operator (%>%) for more readable code

{dplyr}

{tidyr}: Helps you create tidy data

{stringr}: Provide functions for working with strings

{rvest}: Web scrapping

{jsonlite}: JSON parser and generator

**Tidy Data**: Describes a standard way of storing data that is used wherever possible in tidyverse

1 - Every Column is a variable

2 - Every row is an observation

3 - Every cell is a single value

The **dplyr** main verbs (functions)

select(): Pick variables based on their names

filter(): Pick cases based on their values

mutate(): Add new variables as functions of existing ones

arrange(): change the ordering of the rows

summarize(): Reduce multiple values to a single summary

**stringr** verbs

str_detect(x, patterns): Tells you if theres is any match with the pattern

str_count()(x, pattern): Counts the number of patterns

str_subset(x, pattern): Extracts the matching components

str_locate(x, pattern): GIves the position of the match

str_extract(x, pattern): Exttracts the text of the match

str_replace(x, pattern, replacement): Replaces matches with new text

str_split(x, pattern): SPlits up a string into multiple pieces

str_to_lower(), str_to_upper(), str_to_title()

str_pad(): adds spaces to the beginning or end

str_trim(), str_squish()

str_length()

**rvest** package for web scrapping

Designed to work with magrittr

Makes it easy to express common webscrapping tasks

Inspired by libraries like *beautiful soup*

**jronlite** package optimized for the web

It can convert between R objects and JSON without loss of type or information

It is ideal for interacting with web APIs

Key functions: fromJSON() and toJSON()

# The Pipe Operator

```{r}
library(magrittr)
```

To add the pipe operator press ctrl+shift+m. THis is fucking cool!

```{r}
1:10 %>% mean()
```

Make it sexier according to convention

```{r}
1:10 %>% 
  mean()
```

```{r}
1:10 %>% 
  mean() %>% 
  round(digits = 0)
```

```{r}
1:10 %>% 
  mean() %>% 
  floor()
```

```{r}
1:10 %>% 
  mean() %>% 
  floor() %>% 
  print()
```

```{r}
paste("Hello", "World")
```

```{r}
"Hello" %>% 
  paste("World")
```

# Dplyr: The **filter** Verb

```{r}
library(dplyr)
```

```{r}
starwars
```

```{r}
colnames(starwars)
```

```{r}
dim(starwars)
```

```{r}
# Data Frame
starwars["eye_color"]
```

```{r}
# Vector
starwars[["eye_color"]]
```

```{r}
class(starwars["eye_color"])
```

```{r}
starwars["eye_color"] %>% 
  class()
```


```{r}
starwars %>% 
  select(eye_color) %>% 
  unique()
```

```{r}
# Another Way
starwars$eye_color %>% unique()
```

```{r}
starwars %>% 
  filter(eye_color == "black")
```

```{r}
starwars %>% 
  filter(eye_color != "black" & skin_color == "gold")
```

```{r}
starwars %>% 
  filter(eye_color != "black",
         skin_color == "gold",
         height > 100)
```

```{r}
starwars %>% 
  filter(eye_color != "black") %>%
#  filter(skin_color == "gold") %>% 
  filter(height > 100)
```

Dplyr: The **select** Verb

```{r}
which(colnames(starwars) == "height")
```

```{r}
starwars %>% 
  select(2:7)
```

```{r}
starwars %>% 
  select(c(2,3,4,5,6,7))
```

```{r}
which(colnames(starwars) == "hair_color")
```

```{r}
starwars %>% 
  select(c(2,3,5,6,7))
```

```{r}
starwars %>% 
  select(c(7,2,3,5,6))
```

```{r}
# Do the same without the need to supply a vector
starwars %>% 
  select(7,2,3,5,6)
```

In console, write a function and press ctrl + up arrow to see the history of all commands you've written before

```{r}
# No need for index numbers even
# This time we can create a range
starwars %>% 
  select(which(colnames(starwars) == "height"):which(colnames(starwars)== "birth_year"))
```

```{r}
# Do the same by replacing the variable with full stop
starwars %>% 
  select(which(colnames(.) == "height"):which(colnames(.)== "birth_year"))
```

```{r}
# Again same results
starwars %>% 
  select(height:birth_year)
```

```{r}
# Remove hair color from the range
starwars %>% 
  select(height:birth_year, -hair_color)
```

everything function within select

```{r}
starwars %>% 
  select(everything())
```

```{r}
temp <- starwars %>% 
  select(everything())
```


```{r}
library(waldo)
waldo::compare(starwars, temp)
```

```{r}
starwars2 <- starwars %>% 
  select(height:birth_year)
waldo::compare(starwars, starwars2)
```

```{r}
# First pick height to birth, and then everything afterwards
starwars %>% 
  select(height:birth_year, everything())
```

start_with(), end_with(), contains()

```{r}
starwars %>% 
  select(ends_with("color"))
```

```{r}
iris %>% 
  select(Species, starts_with("petal"))
```

```{r}
starwars %>% 
  select(where(is.integer))
```

```{r}
starwars %>% 
  select(where(is.character))
```

```{r}
iris %>% 
  select(ends_with("width"))
```

```{r}
iris %>% 
  select(!ends_with("width"))
```

```{r}
iris %>% 
  select(starts_with("petal") | ends_with("width"))
```

```{r}
starwars %>% 
  select(!c(height, birth_year))
```

# Dplyr: **mutate** Verb

```{r}
library(dplyr)
starwars
```

```{r}
starwars %>% 
  mutate(bmi = mass/(height/100)^2) %>% 
  select(name, height, mass, bmi) %>% 
  mutate(bmi = round(bmi,2))
```

```{r}
# even easier and neater
starwars %>% 
  mutate(bmi = mass/((height/100)^2), bmi = round(bmi,2)) %>% 
  select(name, height, mass, bmi)
```

```{r}
# Temporary columns creation
iris %>%
  select(Species, starts_with("petal")) %>% 
  mutate(Petal.Area = Petal.Length * Petal.Width) %>% 
  filter(Petal.Area > 4 & Petal.Area < 6) %>% 
  select(-Petal.Area)
```

```{r}
starwars %>% 
  select(name, mass) %>% 
  mutate(mass2 = mass*2, mass2_squared = mass2*mass2)
```

```{r}
# Modify multiple columns at once
starwars %>% 
  select(name, homeworld, species) %>% 
  mutate(homeworld = as.factor(homeworld), 
         species = as.factor(species))
```

```{r}
# Make it shorter
starwars %>% 
  select(name, homeworld, species) %>% 
  mutate(across(c(homeworld, species), as.factor))
```

```{r}
# Same, even neater
starwars %>% 
  select(name, homeworld, species) %>% 
  mutate(across(!name, as.factor))
```

# Dplyr: **arrange** Verb

Order rows in ascending/descending criteria

```{r}
# default is ascending. For descending put a - before the column name
starwars %>% 
  arrange(height)
```

```{r}
starwars %>% 
  arrange(desc(height))
```

Nested arrangements

```{r}
mtcars %>% 
  arrange(cyl, disp)
```

```{r}
iris %>% 
  arrange(Sepal.Length, Sepal.Width)
```

```{r}
iris %>% 
  arrange(across())
```

# Dplyr: **summarize** Verb

```{r}
mtcars %>% 
  summarize(mean = mean(disp))
```

```{r}
# Summarize with numbe rof rows being summarized
mtcars %>% 
  summarize(mean = mean(disp), n = n())
```

```{r}
mtcars %>% 
  group_by(cyl) %>% 
  summarize(mean = mean(disp), n = n())
```

```{r}
mtcars %>% 
  group_by(carb) %>% 
  summarize(mean = mean(disp), n = n())
```

```{r}
starwars %>% 
  group_by(species) %>% 
  summarize(count = n(),
            mean_mass = mean(mass, na.rm = TRUE)
            ) %>% 
  filter(count >1, mean_mass > 50) %>% 
  arrange(desc(count))
```

# Dplyr: **tidyr** Verb for Data Pivoting

```{r}
library(tidyr)
```

pivot_longer()

pivot_wider()

Converts a dataset from long to wide, and from wide to long format

```{r}
messy_data <- tibble::tribble(
  ~Student, ~Quiz1, ~Quiz2, ~Test1,
  #-----/-----/--------/-----------/
  "Peter", "F", NA, NA,
  "Bob", "B", "C", "B",
  "John", "A", "A", "B"
)
messy_data
```

This is a wide data format and we wanna convert it into long format

```{r}
pivot_longer(messy_data, c(Quiz1, Quiz2, Test1))
```

```{r}
# Same
tidy_data <- pivot_longer(messy_data, -Student, 
                          names_to = "Assessment", values_to = "Grade")
tidy_data
```

```{r}
messy_data2 <- pivot_wider(tidy_data, 
                           names_from = Assessment, values_from = Grade)
messy_data2
```

```{r}
waldo::compare(messy_data, messy_data2)
```

```{r}
# Funny dataset in tidyr
relig_income
```

Collapse all these into two coulmns: Income range vs counts

```{r}
pivot_longer(relig_income, -religion, names_to = "income", values_to = "count")
```

```{r}
billboard
```

```{r}
billboard_longer <- billboard %>% 
  pivot_longer(cols = starts_with("wk"),
               names_to = "week",
               values_to = "rank")
billboard_longer
```

```{r}
billboard_longer %>% 
  filter(artist == "2 Pac")
```

```{r}
billboard_longer <- billboard %>% 
  pivot_longer(cols = starts_with("wk"),
               names_to = "week",
               values_to = "rank", 
               values_drop_na = TRUE)
billboard_longer
```

```{r}
billboard_longer %>% 
  filter(artist == "2 Pac")
```


```{r}
billboard_longer %>% 
  group_by(track) %>% 
  summarize(highest_rank = min(rank), week = week) %>% 
  arrange(highest_rank)
```

# String Manipulation

```{r}
library(stringr)
```

```{r}
pivot_longer(billboard, 
             starts_with("wk"), 
             names_to = "week", 
             values_to = "rank",
             values_drop_na = TRUE)
```

We wanna split week column and remove the character part

```{r}
x <- c("why", "video", "cross", "extra", "deal", "authority")
# measure the length of each string in the string vector
str_length(x)
```

```{r}
str_sub("Hello", 1, 2)
```

```{r}
str_sub("Hello", 1, 4)
```

```{r}
str_sub("Hello", 3, 4)
```

```{r}
# applies to each string in the vector
str_sub(x, 1, 2)
```

```{r}
pivot_longer(billboard, 
             starts_with("wk"), 
             names_to = "week", 
             values_to = "rank",
             values_drop_na = TRUE) %>% 
  mutate(week = as.integer(str_sub(week, 3, 4)))
```

```{r}
# Same thing done differently
pivot_longer(billboard, 
             starts_with("wk"), 
             names_to = "week", 
             values_to = "rank",
             values_drop_na = TRUE) %>% 
  mutate(week = str_sub(week, 3, 4)) %>% 
  mutate(week = as.integer(week))
```

```{r}
# Find a pattern
str_detect(x, "a")
```

Check the cheatsheet in stringr package from tidyverse

```{r}
# start with
str_detect("Hello World", "^H")
```

```{r}
# end with
str_detect("Hello World", "d$")
```

```{r}
str_extract(x, "ss")
```

```{r}
str_extract(x, "[:alpha:]ss")
```

```{r}
str_extract(x, "[:alpha:]*ss")
```

```{r}
# it picks only the first occurence
string <- "Hello World! Hell yeah!"
str_extract(string, "ll")
```

```{r}
str_extract_all(string, "ll")
```

```{r}
# replaces only the first
str_replace(string, "ll", "x")
```

```{r}
#replaces all
str_replace_all(string, "ll", "x")
```

```{r}
str_split("this, is, a, list", ",") %>% 
  unlist() %>% 
  str_trim()
```

```{r}
str_split("this, is, a, list", ",") %>% 
  unlist() %>% 
  str_squish()
```

```{r}
# adds spaces to reach the length given in the function
str_pad("Hello", 10)
```

```{r}
str_pad("Hello", 10, side = "right")
```

```{r}
str_pad("Hello", 10, side = "both")
```

```{r}
str_pad("Hello", 10, side = "both", pad = "-")
```

# Web Scraping: **rvest**

```{r}
library(rvest)
```

https://www.imdb.com/chart/top/

html tags (ex: <html>, <body>, <div>)
  <html></html>
  <body></body>
  
attributes
  <div id = "" class=:"" styles=""><div/>
  
```{r}
url <- "https://www.imdb.com/chart/top/"

top_movies <- read_html(url)
```

```{r}
top_movies %>% 
  html_nodes("div")
```

```{r}
top_movies %>% 
  html_nodes("div a")
```

```{r}
top_movies %>% 
  html_nodes("tbody tr")
```

```{r}
top_movies %>% 
  html_nodes("tbody tr td.titleColumn")
```

Extract the text within them

```{r}
titles <- top_movies %>% 
  html_nodes("tbody tr td.titleColumn") %>% 
  html_text()
#titles
```

Clean the stuff up

```{r}
# str_split(titles, "\n")
```

```{r}
titles <- titles %>% 
  str_trim() %>% 
  str_split("\n")
```


```{r}
# lapply(titles, function(movie){
#    print(movie)
# })
```

```{r}
lapply(titles, function(movie){
  movie[2]
}) %>%
  str_trim() %>% 
  unlist()
```

```{r}
# merge everything together
titles <- top_movies %>% 
  html_nodes("tbody tr td.titleColumn") %>% 
  html_text() %>% 
  str_trim() %>% 
  str_split("\n") %>% 
  lapply(function(movie) {
    movie[2]
  }) %>% 
  unlist() %>% 
  str_trim()
```

Extract years 

```{r}
years <- top_movies %>% 
  html_nodes("tbody tr td.titleColumn") %>% 
  html_text() %>% 
  str_trim() %>% 
  str_split("\n") %>% 
  lapply(function(movie) {
    movie[3]
  }) %>% 
  unlist() %>% 
  str_trim() %>% 
  str_replace("\\(", "") %>% 
  str_replace("\\)", "") %>% 
  as.integer()
```

Create Ranks

```{r}
ranks <- 1:250
```

```{r}
ratings <- top_movies %>% 
  html_nodes(".imdbRating strong") %>% 
  html_text() %>% 
  as.numeric()
```

Create a dataframe with everything

```{r}
top_movies_tibble <- tibble(
  Rank = ranks,
  Title = titles,
  Year = years,
  Rating = ratings
  )
top_movies_tibble %>% head()
```

```{r}
top_movies_tibble %>% 
  arrange(desc(Year)) %>% 
  head()
```

```{r}
top_movies_tibble %>% 
  group_by(Year) %>% 
  summarize(Average_Rating = mean(Rating)) %>% 
  arrange(desc(Average_Rating))
```

Extract director names by extracting the link of each movie

```{r}
links <- top_movies %>%
  html_nodes("tbody tr td.titleColumn a") %>% 
  html_attr("href") %>% 
  paste0("https://www.imdb.com", .)
links[1]
```

```{r}
links[1] %>% 
  read_html() %>% 
  html_nodes("div.credit_summary_item a") %>% 
  html_text()
```

```{r}
lapply(links[1:10], function(link){
  link %>% 
    read_html() %>% 
    html_node("div.credit_summary_item a") %>% 
    html_text()
}) %>% unlist()
```

# JSON Parsing

```{r}
library(jsonlite)
```

```{r}
mtcars_json <- mtcars %>% 
  toJSON()
```

```{r}
fromJSON(mtcars_json)
```

```{r}
all_equal(mtcars, fromJSON(mtcars_json))
```

```{r}
json <- '["Mario", "Peach", null, "Browser"]'
fromJSON(json)
```

```{r}
fromJSON(json, simplifyVector = FALSE)
```

```{r}
json <- '[
{"Name" : "Mario", "Age": 31, "Occupation" : "Plumber"},
{"Name" : "Peach", "Age": 21, "Occupation" : "Princess"},
{},
{"Name" : "Browser", "Occupation" : "Koopa"}
]'
fromJSON(json)
```

```{r}
my_df <- fromJSON(json)
my_df$ranking <- c(3,1,2,4)
my_df
```

```{r}
toJSON(my_df)
```

```{r}
toJSON(my_df, pretty = TRUE)
```
