---
title: "Part 3 - Intermediate R"
output:
  html_document:
    df_print: paged
---

# Important Functions Related to Packages

install.packages()

installed.packages()

remove.packages()

old.packages(): outdated packages

update.packages()

library()/detach()

Vignettes: Long-form documentation, like a book chapter which shows the reader how to solve a problem. It is accessed by browseVignettes() or vignette()

# Factors

Used to handle categorical variables. The data object is used to categorize data and store it a slevels. They can store both strings and integers and are very important for data analysis and statistical learning.

forcats package from tidyverse provides many tools to deal with factors

fct_reorder(): reordering a factor by another variable

fct_infreq(): reordering a factor by frequency of values

fct_relevel(): changing the order of a factor by hand

fct_lump(): collapse least/most frequent values of factor

# Dates and Time

It is a difficult topic not just in R but all languages

lubridate() makes it easier to create a date/time

Three ways to create a date/time

From a string

from individual date-time components

From an existing date/time object

# The Apply family

apply()/lapply()/sapply()/mapply()/vapply()

The most important is lapply()

Apply a given function to every element of a list and obtains a list as a result

use unlist() to convert any list into a normal flat vector

# Importin/Exporting Data

Tidyverse fir the win

readr: reads rectangular data (.csv. tsv. fwf)

read_csv()

write_csv()

readxl() for getting data out of excel and into R: read_xls()/read_xlsx()

httr() for retrieving webpages

# Relational Operators in R

Operators that compare two terms

```{r}
v = c(2,5.5, 6, 9)
t <- c(8, 2.5, 14, 9)
```

```{r}
v > t
```

```{r}
v < t
```

```{r}
v >= t
```

```{r}
v <= t
```

Find if there is at least one value that fullfills the condition

```{r}
sum(v == t)
```

```{r}
sum(v == t) > 0
```

```{r}
sum(v != t)
```

# Logical Operators

## AND, OR, NOT

```{r}
TRUE & TRUE
```

```{r}
TRUE & FALSE
```

```{r}
TRUE | FALSE
```

```{r}
v1 <- c(TRUE, TRUE, FALSE, FALSE)
v2 <- c(TRUE, FALSE, TRUE, FALSE)
```

```{r}
v1 & v2
```

```{r}
v1 | v2
```

```{r}
!v1
```

```{r}
!v2
```

Coercing numbers into logicals

```{r}
v <- c(3,0, TRUE, 54534)
t <- c(4, 0, FALSE, 3234)
```

```{r}
v&t
```

```{r}
v | t
```

```{r}
!v
```

```{r}
!t
```

# Conditional Statements
## if/else statements

```{r}
if (TRUE) {
  print("Hello, World!")
}
```

```{r}
var <- 10
var > 10
```

```{r}
if ( var > 10) {
  print("Var is Greater than 10")
}
```

```{r}
if ( var <= 10) {
  print("Var is Smaller than or Equal to 10")
  }
```

```{r}
if ( var > 10) {
  print("Var is Greater than 10")
} else {
  print("Var is less than or equal to 10")
}
```

```{r}
var <- 11
var2 <- "Hi"
if ( var > 10 & var2 == "Hi"){
  print("That works!")
}
```

```{r}
var = 3
if (var > 10) {
  print("Var is greater than 10")
} else if (var == 10) {
  print("Var is equal to 10")
} else {
  print("Var is smaller than 10")
}
print("Hello World!")
```

# Working With Loops

```{r}
v <- LETTERS[1:4]
for (i in v) {
  print(i)
}
```

Do the same with lapply

```{r}
results <- lapply(v, print)
unlist(results)
```

```{r}
result <- sapply(v, print)
unlist(result)
```

# Working With Functions

```{r}
seq(5, 15, by = 4)
```

```{r}
sum(1:10)
```

Function without any argument

```{r}
func_x <- function() {
  print("Hello World")
}
func_x()
```

```{r}
f_square <- function() {
  (1:10)^2
}
f_square()
```

Calling the function definition

```{r}
f_square
```

```{r}
f_square <- function(a = 1:10) {
  (a)^2
}
f_square()
```

```{r}
f_square(10)
```

```{r}
f_square(1:5)
```

# Packages Installment and Management

```{r}
# install.packages(c("skimr", "corrr"))
library(skimr)
library(corrr)
```

```{r}
#installed.packages()
```

```{r}
old.packages()
```

```{r}
help(ggplot2)
browseVignettes(package = "corrr")
vignette("using-corrr")
```

# Working With Factors

Lets start with a simple string vector

```{r}
data <- c("East", "West", "East", "North", "North", "East", 
          "West", "West", "West", "East", "North", "West")
is.factor(data)
class(data)
```

```{r}
unique(data)
```

factor() and as.factor() functions do the same thing

```{r}
data_factor <- factor(data)
data_factor
```

```{r}
levels(data_factor)
```

```{r}
as.integer(data_factor)
```

Specify the order as you wish

```{r}
sort(data_factor)
```

```{r}
data_factor_new <- factor(data_factor, levels = c("East", "West", "North"))
data_factor_new
```

```{r}
is.integer(data_factor) == is.integer(data_factor_new)
```

```{r}
sort(data_factor_new)
```

```{r}
data_factor_complete <- factor(data_factor_new, levels = c("North", "East", "South", "West"))
data_factor_complete
```

```{r}
as.integer(data_factor_complete)
```

```{r}
table(data_factor_complete)
```

```{r}
table(data_factor)
```

```{r}
sort(data_factor_complete)
```

```{r}
data_factor_new_2 <- factor(data_factor_complete, levels = c("West", "East", "North", "South"))
table(data_factor_new_2)
```


```{r}
library(forcats)
fct_infreq(data_factor_complete)
```

```{r}
fct_relevel(data_factor_complete, "North")
```

# Dates and Time

We use lubridate package from tidyverse to deal with time and date

```{r}
library(lubridate)
```

```{r}
today()
```

```{r}
class(today())
```

```{r}
now()
```

```{r}
class(now())
```

Create a date from a string

```{r}
string <- "2017-01-01"
date <- ymd(string)
class(date)
```

```{r}
mdy("01-02-2017")
```

```{r}
mdy("January 31st, 2017")
```

```{r}
date2 <- mdy("January 31st, 2017")
class(date2)
```

```{r}
dmy("31-Jan-2017")
```

Creating date object from numbers

```{r}
ymd(20170131)
```

```{r}
date_time <- ymd_hms("2017-01-31 20:11:59")
class(date_time)
```

```{r}
ymd_hm("2017/01/31 20:11")
```

```{r}
mdy_hm("01/31/2018 08:22")
```

# Functional Programming

We wont cover everything as this is a huge topic, but will concentrate mostly on lapply()

```{r}
# The function below is the same as lapply()
# lapply(c(10,10,10,10,10,10,10,10,10,10), runif)
values <- lapply(rep(10, 10), runif)
values
```

```{r}
result <- runif(10)
mean(result)
```

```{r}
means <- lapply(values, mean)
means
```

As lapply() always returns a list, lets flatten it into a vector of values

```{r}
unlist(means)
```

```{r}
unlist(lapply(values, mean))
```

```{r}
nchar("Hey")
```

```{r}
nchar("Hello World")
```

```{r}
car_names <- rownames(mtcars)
unlist(lapply(car_names, nchar))
```

```{r}
car_names_length <- unlist(lapply(car_names, nchar))
car_names[4]
car_names_length[4]
```

```{r}
car_nm <- "Hornet 4 Drive"
which(car_names == car_nm)
```

```{r}
which(car_names == "Fiat X1-9")
```

```{r}
which(rownames(mtcars) == "Hornet Sportabout")
```

```{r}
which(colnames(mtcars) == "am")
```

```{r}
mtcars[5,9]
```

```{r}
random_function <- function(name) {
  
  if (mtcars[name,9] == 0) # if autotmatic (=0)
    mtcars[name,1] 
  else
    mtcars[name,1]^2 # if manual (=0)
}
random_function("Hornet Sportabout")
```

```{r}
unlist(lapply(rownames(mtcars), random_function))
```

```{r}
unlist(lapply(rownames(mtcars), random_function <- function(name) {
  
  if (mtcars[name,9] == 0) # if autotmatic (=0)
    mtcars[name,1] 
  else
    mtcars[name,1]^2 # if manual (=0)
}))
```

# Data Import and Export

Using tidyverse alternatives that have no discrepencies. We use readr. The main functions are read_csv() and write_csv()

```{r}
library(readr)
read_csv("test.csv")
```

```{r}
?read_csv
```

```{r}
read_csv("test.csv", col_names = FALSE)
```

```{r}
df <- read_csv("test.csv")
df
```

```{r}
iris
```

```{r}
write_csv(iris, file = "iris.csv")
```

```{r}
read_csv("iris.csv")
```

```{r}
library(readxl)
excel_import <- read_xlsx("test.xlsx")
excel_import
```

read webpages

```{r}
library(httr)
r <- GET("tidyverse.org")
# httr::GET("tidyverse.org") #Same
r$url
```

```{r}
status_code(r)
```

```{r}
content(r)
```

# Working With DataBases

Relational Databases: SQL

DBI package

```{r}
library(DBI)
```

```{r}
con <- dbConnect(RSQLite::SQLite(), dbname = ":memory:")
```

```{r}
dbListTables(con)
```

Create a table within SQL database

```{r}
dbWriteTable(con, "mtcars", mtcars)
```

```{r}
dbListTables(con)
```

```{r}
dbListFields(con, "mtcars")
```

```{r}
table <- dbReadTable(con, "mtcars")
table
```

```{r}
res <- dbSendQuery(con, "SELECT * FROM mtcars WHERE cyl = 4")
dbFetch(res)
```

Remove the results 

```{r}
dbClearResult(res)
```

Closing the connection whenever you need to

```{r}
dbDisconnect(con)
```
