---
title: "Part 7 - Linear Regression Model"
author: "Ilir_Sheraj"
date: "7/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Tidymodels steps**

1 - Split data {rsample}

2 - Prepare recipe {recipes}

3 - Specify Model {parsnip}

4 - Tune Hyperparameters {tune}

5 - Fit Model {parsnip}

6 - Analyze Model {broom}

7 - Predict {parsnip}

8 - Interpret Results {yardstick}

# Train a Linear Regression Model

```{r}
library(dplyr)
library(ggplot2)
library(corrr)
library(rsample)
library(recipes)
library(parsnip)
```

```{r}
diamonds
```

To select only numerical variables 
```{r}
diamonds %>% 
  select(where(~!is.factor(.x))) %>% 
  corrr::correlate() %>% 
  corrr::fashion()
```

```{r}
diamonds %>% 
  select(where(~!is.factor(.x))) %>% 
  corrr::correlate() %>% 
  corrr::network_plot(min_cor = 0.1)
```

Select by correlation

```{r}
diamonds %>% 
  select(where(~!is.factor(.x))) %>% 
  correlate() %>% 
  focus(price) %>% 
  fashion()
```

Automate the pipeline so that if any other numerical variable is added to the data, it will be automatically pulled

```{r}
diamonds %>% 
  select(where(~!is.factor(.x))) %>% 
  correlate() %>% 
  focus(price) %>% 
  arrange(desc(price)) %>% 
  filter(price > 0.5) %>% 
  pull(term)
```

```{r}
highly_corr_variables <- diamonds %>% 
  select(where(~!is.factor(.x))) %>% 
  correlate() %>% 
  focus(price) %>% 
  arrange(desc(price)) %>% 
  filter(price > 0.5) %>% 
  pull(term)
highly_corr_variables
```

## Build a Linear Model with all the data

```{r}
data_all_split <- initial_split(diamonds, strata = price)
data_all_train <- training(data_all_split)
data_all_test <- testing(data_all_split)
```


```{r}
model_all <- lm(price ~ ., data = diamonds)
summary(model_all)
```

```{r}
broom::tidy(model_all) %>% 
  arrange(p.value)
```

## Select only numerical values

```{r}
data <- diamonds %>% 
  select(all_of(highly_corr_variables), price)
data
```

Default is 75% to 25%

```{r}
data_split <- initial_split(data, strata = price)
data_train <- training(data_split)
data_test <- testing(data_split)
```

```{r}
model <- lm(price ~ ., data = data_train)
```

```{r}
summary(model)
```

```{r}
broom::tidy(model)
```

## Tidy Models Data

```{r}
data_tidy_split <- initial_split(diamonds, strata = price)
data_tidy_train <- training(data_tidy_split)
data_tidy_test <- testing(data_tidy_split)
```

### Preprocessing

```{r}
data_rec <- data_tidy_train %>% 
  recipe(price ~., data_tidy_train) %>% 
  step_dummy(all_nominal()) %>% 
  step_normalize(all_numeric(), - all_outcomes()) %>% 
  prep()

juiced <- juice(data_rec)
```

```{r}
lm_model <- parsnip::linear_reg() %>% 
  set_engine("lm") %>% 
  set_mode("regression")
```

```{r}
lm_fit <- fit(lm_model, price ~ ., data = juiced)
lm_fit
```

```{r}
glance(lm_fit)
```

```{r}
tidy(lm_fit)
```


```{r}
results_train <- lm_fit %>% 
  predict(new_data = juiced) %>% 
  mutate(truth = data_tidy_train$price)
```

```{r}
results_test <- lm_fit %>% 
  predict(new_data = bake(data_rec, data_tidy_test)) %>% 
  mutate(truth = data_tidy_test$price)
```

```{r}
results <- results_train %>% 
  mutate(type = "train") %>% 
  bind_rows(
    results_test %>% 
      mutate(type = "test")
  )
```

```{r}
results %>% group_by(type) %>% 
  yardstick::rmse(truth, .pred)
```

```{r}
ggplot(results, aes(truth, .pred)) +
  geom_point() +
  geom_abline(color = "red", size = 2) +
  coord_fixed()
```

