---
title: "Part 6 - EDA"
author: "Ilir_Sheraj"
date: "7/25/2021"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tibble)
library(dplyr)
```

```{r}
diamonds
```

```{r}
glimpse(diamonds)
```

```{r}
summary(diamonds)
```

There is not a single NA value in hte dataset

```{r}
str(diamonds)
```

```{r}
dim(diamonds)
```

```{r}
skimr::skim(diamonds)
```

```{r}
head(diamonds)
```

```{r}
tail(diamonds)
```

## Lets Visualize the Distributions
### Categorical Variables

```{r}
ggplot(diamonds, aes(cut)) +
  geom_bar()
```

```{r}
diamonds %>% count(cut)
```

```{r}
table(diamonds$cut)
```

```{r}
unique(diamonds$cut)
```

```{r}
levels(diamonds$cut)
```

```{r}
levels(diamonds$clarity)
```

```{r}
table(diamonds$clarity, diamonds$cut)
```

Do the same but order

```{r}
diamonds %>% 
  count(cut, clarity) %>% 
  arrange(n)
```

```{r}
diamonds %>% 
  count(cut, clarity) %>% 
  arrange(desc(n))
```

```{r}
ggplot(diamonds, aes(x = cut, fill = clarity)) +
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45))
```

```{r}
ggplot(diamonds, aes(x = cut, fill = clarity)) +
  geom_bar(position = "dodge") +
  coord_flip()
```

```{r}
ggplot(diamonds, aes(x = cut, fill = clarity)) +
  geom_bar(position = "fill") +
  ylab("Proportion")
```

```{r}
table(diamonds$cut, diamonds$clarity)
```

To calculate the percentages of each

```{r}
prop.table(table(diamonds$cut, diamonds$clarity))
```

```{r}
prop.table(table(diamonds$cut, diamonds$clarity)) %>% sum()
```

```{r}
ggplot(diamonds, aes(x = cut)) +
  geom_bar() +
  theme(axis.text.x = element_text(angle = 45)) +
  facet_wrap(~clarity)
```

```{r}
round(prop.table(table(diamonds$cut, diamonds$clarity))*100, 2)
```

## Working with Continous Variables

```{r}
ggplot(diamonds, aes(x = carat)) +
  geom_histogram(binwidth = 0.5)
```

See the ranges with different intervals

```{r}
diamonds %>% 
  count(cut_width(carat, 0.5))
```

```{r}
smaller <- diamonds %>% 
  filter(carat < 3)
ggplot(smaller, aes(x = carat)) +
  geom_histogram(binwidth = 0.1)
```

```{r}
ggplot(smaller, aes(carat, color = cut)) +
  geom_freqpoly(binwidth = 0.1)
```

```{r}
ggplot(smaller, aes(carat)) +
  geom_histogram(binwidth = 0.01)
```

```{r}
ggplot(diamonds, aes(y)) +
  geom_histogram(binwidth = 0.5) +
  coord_cartesian(ylim = c(0,50))
```

Zero values are outliers and seem to be mistakes because its impossible to have diamonsw without dimensions

```{r}
diamonds %>% filter(y == 0)
```

```{r}
diamonds %>% filter(y > 20)
```

```{r}
# Outliers
diamonds %>% 
  filter(y < 3 | y > 20) %>% 
  select(price, x, y, z) %>% 
  arrange(y)
```

## Missing Values

```{r}
diamonds %>% 
  filter(between(y, 3, 20)) %>% 
  ggplot(aes(y)) + 
  geom_histogram(binwidth = 0.5)
```

```{r}
diamonds %>% 
  mutate(y = ifelse(y < 3 | y > 20, NA, y)) %>% 
  ggplot(aes(x, y)) +
  geom_point()
```

```{r}
diamonds %>% 
  ggplot(aes(x, y)) +
  geom_point()
```

## Covariation
### Categorical vs Continuous

```{r}
diamonds %>% 
  ggplot(aes(price)) +
  geom_freqpoly(aes(color = cut), binwidth = 500)
```

```{r}
diamonds %>% 
  ggplot(aes(cut)) +
  geom_bar()
```

### Density Plot

```{r}
diamonds %>% 
  ggplot(aes(price, y = ..density..)) +
  geom_freqpoly(aes(color = cut, binwidth = 500))
```

```{r}
diamonds %>% 
  ggplot(aes(cut, price)) +
  geom_boxplot() +
  coord_flip()
```

## Categorical vs Categorical

```{r}
diamonds %>% 
  ggplot(aes(cut, color)) +
  geom_count()
```

```{r}
diamonds %>% 
  count(color, cut) %>% 
  ggplot(aes(color, cut)) +
  geom_tile(aes(fill = n))
```

```{r}
diamonds %>% 
  ggplot() +
  geom_point(aes(x = carat, y = price), alpha = 1/100)
```

```{r}
smaller %>% 
  ggplot() +
  geom_bin2d(aes(carat, price))
```

